\documentclass{article}

\usepackage[T1]{fontenc}
\usepackage[osf]{libertine}
\usepackage[scaled=0.8]{beramono}
\usepackage[margin=1.5in]{geometry}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{bm}

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}

\usepackage{titlesec}
\titlespacing{\section}{0pt}{10pt plus 2pt minus 2pt}{0pt plus 2pt minus 0pt}
\titlespacing{\subsection}{0pt}{5pt plus 2pt minus 2pt}{0pt plus 2pt minus 0pt}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\newcommand{\acro}[1]{\textsc{\MakeLowercase{#1}}}
\newcommand{\given}{\mid}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\data}{\mc{D}}
\newcommand{\intd}[1]{\,\mathrm{d}{#1}}
\newcommand{\mat}[1]{\bm{\mathrm{#1}}}
\renewcommand{\vec}[1]{\bm{\mathrm{#1}}}
\newcommand{\trans}{^\top}

\begin{document}

{\large \textbf{CSE 515T (Spring 2018) Assignment 3}} \\
Due Monday, 19 March 2018 \\

\begin{enumerate}

\item
  (Sampling from a multivariate Gaussian distribution).  Assume you can sample
  from a univariate standard normal distribution $\mc{N}(0, 1^2)$.  In the last
  assignment, this allowed you to sample trivially from a multivariate standard
  normal distribution $\mc{N}(\vec{0}, \mat{I})$.  Use the closure of the
  Gaussian distribution to affine transformations to derive a procedure to
  sample from an arbitrary multivariate Gaussian distribution
  \[
    p(\vec{x}) = \mc{N}(\vec{x}; \vec{\mu}, \mat{\Sigma})
  \]
  using an oracle that can draw a sample from the univariate standard normal
  distribution.

\item
  (Sampling from a Gaussian process).
  Consider the squared exponential covariance function
  \begin{equation}
    \label{sqdexp}
    K(x, x'; \lambda, \ell)
    =
    \lambda^2
    \exp\biggl(
    -\frac{\lVert x - x' \rVert^2}{2\ell^2}
    \biggr).
  \end{equation}
  Take a dense grid (somewhere around 1\,000 points) in the domain $x \in [-4,
    4]$.  For $(\lambda, \ell) = (1, 1)$, draw five samples from a Gaussian
  process prior with mean function $\mu(x) = 0$ and this covariance.  Plot your
  samples.  Repeat with $(\lambda, \ell) = (2, 1)$ and $(\lambda, \ell) = (1,
  2)$.  What changed?

  Note: you may encounter numerical problems here.  The most common issues are
  that the sample covariance $\mat{K}$ is not quite symmetric or that it is not
  quite positive definite.  To ensure positive definiteness, you may need to add
  a small constant (on the order of $10^{-6}$ perhaps) to the diagonal to the
  covariance matrix $\mat{K}$ in case of numerical instability.  If the matrix
  is not exactly symmetric, a simple trick to symmetrize it is to take
  $\frac{1}{2}(\mat{K} + \mat{K}\trans)$.

  Repeat this sampling procedure with the following covariance functions:
  \begin{align*}
    K(x, x')
    &=
    \exp\bigl(\lvert x - x' \rvert\bigr)
    \\
    K(x, x')
    &=
    \bigl(1 + \lvert x - x' \rvert^2\bigr).
  \end{align*}

\item
  (Gaussian process regression).
  Consider the following data:
  \begin{align*}
    \vec{x}
    &=
    [-2.26, -1.31, -0.43, 0.32, 0.34, 0.54, 0.86, 1.83, 2.77, 3.58]\trans; \\
    \vec{y}
    &=
    [1.03, 0.70, -0.68, -1.36, -1.74, -1.01, 0.24, 1.55, 1.68, 1.53]\trans.
  \end{align*}
  Fix the observation noise variance at $\sigma^2 = 0.5^2$.

  \begin{itemize}
  \item
    Examining a scatter plot of the data, guess which values of $(\lambda,
    \ell)$ in the above covariance \eqref{sqdexp} (if any) might explain this
    data well.
  \item
    Perform Gaussian process regression for these data on the interval $x_\ast
    \in [-4, 4]$ using the squared exponential covariance \eqref{sqdexp} for the
    same set of hyperparameters $(\lambda, \ell)$ above.  Plot the posterior
    mean and the pointwise 95\% credible interval for each.  Which predictions
    look the best?
  \item
    Visualize the model evidence $p(\vec{y} \given \vec{x}, \lambda, \ell,
    \sigma^2)$ as a function of $(\lambda, \ell)$.  You can choose to make, for
    example, a heatmap or a contour plot of this function.  Probably it will be
    a good idea to compute and plot the logarithm of the model evidence rather
    than the model evidence directly.
  \end{itemize}
\end{enumerate}

\end{document}
